World Heritage Sites on Wikipedia: Cultural Heritage Activism in a Context of Constrained Agency

### 2021-03-31 analysis/paper/paper.Rmd
update rnaturalearth

Of the `r nrow(wh_unesco)` cultural sites on the World Heritage List at the time of writing, we found Wikipedia articles for `r nrow(some_page_variables)`. As a group, the basic details of content for Wikipedia articles for CS-WHL differ little from a sample of 10,000 random Wikipedia articles (Figure \@ref(fig:wh-wiki-content-and-consumption)). The scholarly nature of the articles, measured by the number of sources cited in the reference list per thousand words in the article body, have similarly distributions for CS-WHL articles and random articles. The number of Wikilinks out from the target article to other Wikipedia articles are also similarly distributed for CS-WHL articles and random articles. The total number of words in a CS-WHL article is typically much higher than a random article, indicating that they receive more generative effort from editors than other articles.

@@

```{r wh-wiki-consumption, fig.cap="Consumption of Wikipedia articles about CS-WHL. The density plots on the left show basic characterstics of consumption. The right shows a scatterplot of consumption metrics for Wikipedia articles about CS-WHL, with labels on the most intensively consumed articles", fig.width=15}
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_articles_basic_consumption.png"))
```

The basic metrics of consumption of CS-WHL articles show substantial differences from our sample of random articles (Figure \@ref(fig:wh-wiki-content-and-consumption)). We measure consumption by counting the total number of views of the article over the 100 days prior to our data collection date, and the the number of Wikilinks from other articles into the target article. Wikipedia article view counts are popular widely-used measures of cultural interest or salience [@roll2016using; @cao2020analysis; @mciver2014wikipedia]. Wikilinks from other articles are a measure of the centrality of an article, if many other articles link to it, then the article is well-integrated into the encyclopedia and viewed as important for supporting information presented in other articles. CS-WHL articles are typically viewed far more frequently than other Wikipedia articles, reflecting high consumption by internet users generally. They are also much more often linked to by other Wikipedia articles than our random sample of other articles, indicating consumption by other Wikipedia articles and Wikipedia users in their editing work (Figure \@ref(fig:wh-wiki-content-and-consumption)). This indicates that consumption of CS-WHL articles is generally high, relative to other articles, and confirms our assertion of Wikipedia as an important source of heritage information. But is how is attention distributed across all CS-WHL articles, and how does it relate to sites with conflicts and tensions?

@@

We can see an answer to this question in the scatterplot on the lower right of Figure \@ref(fig:wh-wiki-content-and-consumption), which shows the values of inward Wikilinks, article views, and article word count for all CS-WHL articles. The labeled points in the upper right quadrant are the articles that receive most of the attention. As noted above for sites with high proportions of reverts and edits, highly iconic sites in the Western canon of culture history are also the majority of sites that are highly consumed. Of the CS-WHL notable for conflict and tension, only Timbuktu is visible among these highly popular articles. It is also the only site on the List of World Heritage in Danger that is among these highly popular articles. We reviewed the talk page for the Timbuktu article to determine if the attention received by the article might relate to the issues leading to armed conflict at the site. Of the `r timbuktu_talk_page_word_count` words on the talk page, only 168 are on the topic of conflict and destruction, with the editors discussing how to describe the scale of the damage to the temples. The majority of talk page content for Timbuktu is about filling in missing detail, suggestions for or notifications of minor corrections. This is also the case for the talk page for Auschwitz, another popular CS-WHL article with six archived talk pages including over 100,000 words. Although Wikipedia articles in other languages on Auschwitz have different points of emphasis [@wolniewicz2016framing], the majority of the discussion among editors on the talk pages is technical and precise. Provocative comments on the English language Auschwitz talk pages, for example by editors denying the Holocaust, usually end after brief exchanges with other editors requesting credible sources. The rarity of conflict on the Auschwitz talk page can be contrasted with the talk page for the Holocaust, where @pfanzelter2015crossroads found abundant evidence of conflict among editors. Generally, our results show that conflict and tension at a CS-WHL is not highly correlated with how much attention an article receives.

>>>>

Of the `r nrow(wh_unesco)` cultural sites on the World Heritage List at the time of writing, we found Wikipedia articles for `r nrow(some_page_variables)`. As a group, the basic details of content for Wikipedia articles for CS-WHL differ little from a sample of 10,000 random Wikipedia articles (Figure \@ref(fig:wh-wiki-content)). The scholarly nature of the articles, measured by the number of sources cited in the reference list per thousand words in the article body, have similarly distributions for CS-WHL articles and random articles. The number of Wikilinks out from the target article to other Wikipedia articles are also similarly distributed for CS-WHL articles and random articles. The total number of words in a CS-WHL article is typically much higher than a random article, indicating that they receive more generative effort from editors than other articles.

@@

```{r wh-wiki-consumption, fig.cap="Consumption of Wikipedia articles about CS-WHL. The density plots on the top show basic characterstics of consumption for Wikipedia articles about CS-WHL (yellow) compared to 10,000 random Wikipedia articles (grey). The scatterplot at the bottom shows consumption metrics for Wikipedia articles about CS-WHL, with labels on the most intensively consumed articles", fig.width=15}
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_articles_basic_consumption.png"))
```

The basic metrics of consumption of CS-WHL articles show substantial differences from our sample of random articles (Figure \@ref(fig:wh-wiki-consumption)). We measure consumption by counting the total number of views of the article over the 100 days prior to our data collection date, and the the number of Wikilinks from other articles into the target article. Wikipedia article view counts are popular widely-used measures of cultural interest or salience [@roll2016using; @cao2020analysis; @mciver2014wikipedia]. Wikilinks from other articles are a measure of the centrality of an article, if many other articles link to it, then the article is well-integrated into the encyclopedia and viewed as important for supporting information presented in other articles. CS-WHL articles are typically viewed far more frequently than other Wikipedia articles, reflecting high consumption by internet users generally. They are also much more often linked to by other Wikipedia articles than our random sample of other articles, indicating consumption by other Wikipedia articles and Wikipedia users in their editing work (Figure \@ref(fig:wh-wiki-consumption)). This indicates that consumption of CS-WHL articles is generally high, relative to other articles, and confirms our assertion of Wikipedia as an important source of heritage information. But is how is attention distributed across all CS-WHL articles, and how does it relate to sites with conflicts and tensions?

@@

We can see an answer to this question in the scatterplot on the lower part of Figure \@ref(fig:wh-wiki-consumption), which shows the values of inward Wikilinks, article views, and article word count for all CS-WHL articles. The labeled points in the upper right quadrant are the articles that receive most of the attention. As noted above for sites with high proportions of reverts and edits, highly iconic sites in the Western canon of culture history are also the majority of sites that are highly consumed. Of the CS-WHL notable for conflict and tension, only Timbuktu is visible among these highly popular articles. It is also the only site on the List of World Heritage in Danger that is among these highly popular articles. We reviewed the talk page for the Timbuktu article to determine if the attention received by the article might relate to the issues leading to armed conflict at the site. Of the `r timbuktu_talk_page_word_count` words on the talk page, only 168 are on the topic of conflict and destruction, with the editors discussing how to describe the scale of the damage to the temples. The majority of talk page content for Timbuktu is about filling in missing detail, suggestions for or notifications of minor corrections. This is also the case for the talk page for Auschwitz, another popular CS-WHL article with six archived talk pages including over 100,000 words. Although Wikipedia articles in other languages on Auschwitz have different points of emphasis [@wolniewicz2016framing], the majority of the discussion among editors on the talk pages is technical and precise. Provocative comments on the English language Auschwitz talk pages, for example by editors denying the Holocaust, usually end after brief exchanges with other editors requesting credible sources. The rarity of conflict on the Auschwitz talk page can be contrasted with the talk page for the Holocaust, where @pfanzelter2015crossroads found abundant evidence of conflict among editors. Generally, our results show that conflict and tension at a CS-WHL is not highly correlated with how much attention an article receives.
